Renewal and stability in populations structured by remaining years of life

### 2014-09-04 ThanoRepro/ThanoRepro.tex
lots of edits. new appendix for optimization method. more work on matrix stuff, but still not perfectly equal...

\FloatBarrier
\section*{The thanatological renewal model}
For the renewal model that follows, we are most interested in using the kind
of fertility rates shown in Figure~\ref{fig:FySpaghetti}, and we refer to a
unisex population based on female vital rates. In the following, I
explain describe the model in four different ways: as a continuous
model, as a matrix model, with a lifecylcle graph, and intuitively by
appealing to how population pyramids and leaves (described earlier) change over
time. Assuming constant vital rates, the births for the present year are given by
\begin{align}
B(t) =& \int _0^\infty P(y,t)f^\star(y) \dd y = \int _0^\infty P(a,t)f(a) \dd a
\quad\text{,} \intertext{where $a$ indexes chronological age, $y$ indexes
thanatological age, $P(a), P(y)$ are population counts, and $f^\star(y), f(a)$ are
exact specific fertility probabilities (rates), which here are best thought of
as the fertility rates of females born to females. The thanatological integral can
be broken down back in terms of chronological age:} =& \int_{y=0}^\infty
\int_{a=0}^\infty f^\star(y)\frac{P(a,t)d(a+y)}{l(a)} \dd a \dd y \quad\text{,}
\intertext{where $d(a)$ is the continuous lifetable death distribution with
radix of 1, so $d(a)/l(0)$. We can relate the present population to past births
with $P(a,t) = B(t-a)l(a)$:}
\label{eq:ergostep1}
=& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) B(t-a)d(a+y)\dd a \dd y
\quad\text{.} \intertext{Eventually-- usually a bit quicker than is the case for
chronological age-- strong ergodicity will assert itself, and $B(t)$ will be related to
$B(t-a)$ according to a constant factor $e^{ra}$, where $r$ is the familiar
intrinsic rate of growth:}
\label{eq:ergostep2}
=& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) B(t)e^{-ra}d(a+y)\dd a \dd y
\quad\text{.}
\intertext{Divide out $B(t)$ to get back a familiar-looking renewal equation:}
\label{eq:thanoren}
1 =& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) d(a+y)e^{-ra}\dd a \dd y
\quad\text{.}
\end{align}
Now compare this to Lotka's chronological formulation
\begin{equation}
\label{eq:lotka}
1 = \int_{a=0}^\infty f(a)l(a)e^{-ra}\dd a \quad\text{,}
\end{equation}
and note that these are really quite similar, since $\int _{y=0}^\infty
d(a+y)\dd a = l(a)$. For intuition, notice that $l(a)$ is here split
up into pieces of $d(a)$, and imagine a 2D surface of these, where one axis is chronological age and the
other axis is thanatological age. For the chronological case \eqref{eq:lotka},
we multiply chronological age-specific fertility rates over one margin, and for
the thanatological case over the other margin. There is a strong parallel here
with the case of two-sex age-specific fertility rates. As in the case of
divergence between male and female single-sex models, there will always be divergence between
single-sex models under chronological versus thanatological age, unless the
starting population is stable. Even though the modeled population stocks are in
a way commensurable, the rates used here, $f(a)$ versus $f^\star(y)$ are
calculated on the basis of differently distributed denominators $E(a)$ versus $E(y)$.

@@

Once one finds $r$ from \eqref{eq:thanoren}, other familiar stable population
parameters can be calculated. For instance, we may calculate the mean
thanatological generation time, $T^y$:
\begin{equation}
\label{eq:Ty}
 T^y =  \frac{\int _{y=0}^\infty \int _{a=y}^\infty y e^{-ra} d(a) f^\star(y) \dd a
\dd y}{\int _{y=0}^\infty \int _{a=y}^\infty e^{-ra} d(a) f^\star(y) \dd a \dd
y} \quad \text{.}
\end{equation}
Literally, this is the mean of the remaining lifespans of new mothers in the
stable population. The net reproduction rate, $R_0$ is related by, e.g.,
\begin{equation}
\label{eq:R0fromTy}
R_0 = e^{r T^y}
\end{equation}
The birth rate, $b$, is given by
\begin{equation}
\label{eq:eybrate}
b = \frac{1}{\int _{y=0}^\infty \int _{a=y}^\infty e^{-ra} d(a) \dd a
\quad\text{.} \dd y}
\end{equation}
The stable thanatological age structure, $c_y$, is the
proportion of the stable population with remaining years to live $y$:
\begin{equation}
\label{eq:cy}
c(y) = b \int _{a=y}^\infty e^{-ra} d(a) \dd a \quad\text{.}
\end{equation}
We already knew that when $r = 0$ and $y = a$, that $c(a)=c(y)$
\citep{vaupel2009life}, and now we can arrive at the same conclusion directly
by equating the chronological and thanatological renewal equations. For
small non-zero magnitude values of $r$ and survival patterns typical of humans, it is also the
case that $c(y|r)\approx c(a|-r)$. The relationship is asymptotically exact as
$r$ approaches zero. Other findings regarding stable age structure will be
presented in a future version of this paper. Further stable population
quantities may be estimated by similarly translating the various common definitions (e.g.,
in the glossary of \citet{coale1972growth}) to the present perspective. We will
focus on the main model rather than on these.

The thanatological renewal model is coherent, but these ideas may take some time
to gestate before accepting that it is also a valid description of how
populations renew. What should one imagine under the model of thanatological
population renewal? A useful mnemonic bases itself on Figures~\ref{fig:USdecomp}
and \ref{fig:USrecomp}. In the chronological age-structured model, new
generations appear at the bottom of the pyramid, and move up one rung per year.
All age-classes are subject to attrition, which is spread out over ages and not
readily visible in the pyramid. In the thanatological \textit{leaf}, each birth cohort increments to the population over the whole range of thanatological age
according to $d(a)$, as seen in \eqref{eq:thanoren}, becoming the
shaded layers seen in Figure~\ref{fig:USrecomp}. Each horizontal step is a death
cohort, and these move one step down the pyramid each year without any decrement
(indeed incrementing due to births) until reaching the very bottom. In short,
the locations of increment and decrement, and the direction of movement are all
switched. The chronological and thanatological renewal models are almost
perfectly opposite descriptions of the same two processes.

@@

\begin{appendices}
\section{Unique solution for thanatological $r$}
\label{app:A}
The solution for the intrinsic growth rate, $r$, is unique for the case of the
thanatological renewal model, \eqref{eq:thanoren}, and can be proven so in
essentially the same fashion as those in existence for the Lotka-Euler model,
\eqref{eq:lotka}. This little proof follows that given in
\citet{pressat1973analyse}. Define a convenience function, $I(r)$, for the
integrand of \eqref{eq:thanoren} for a given $r$ and fixed $f^\star(y)$ and $d(a)$:

\begin{equation}
I(r) = \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) d(a+y)e^{-ra}\dd a \dd y
\end{equation}
Since the death distribution function, $d(a)$, and fertility
function, $f^\star(y)$, are continuous and non-negative, $\lim_{r \to +\infty} I(r)
= 0$ and $\lim_{r \to -\infty} I(r)= \infty$. If $r_2 > r_1$, then $I(r_1) >
I(r_2)$. $I()$ is therefore a continuous and monotonically decreasing function
of $r$ with boundaries that include the value 1 of \eqref{eq:thanoren}, and
 necessarily only obtain this value once.

\section{Equivalence in stability}
\label{app:B}
This appendix contains a proof that the thanatological renewal equation implies
the same intrinsic growth rate, $r$, as Lotka's chronological renewal model if
the starting population is stable. This is identical to claiming that the right
side of the rewal equation is equal given the same $r$. Since there is a unique
real solution, demonstrated in the previous appendix, then it is sufficient to
show that the two integral equations are equal under these conditions. The basic
relationship is:
\begin{align}
\int _{a=0}^\infty l(a)e^{-ra}f(a) \dd a &= \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra}f^\star (y) \dd a \dd y
\intertext{when}
f^\star(y) &= \frac{b^\star(y)}{c^\star(y)} \\
b^\star (y) &= \int_{a=0}^\infty b(a) \mu(a+y))\frac{l(a+y)}{l(a)} \dd a \\
c^\star (y) &= \int_{a=0}^\infty c(a) \mu(a+y))\frac{l(a+y)}{l(a)} \dd a
\intertext{Note that:}
c(a) &= \frac{l(a)r^{-ra}}{\int l(a) e^{-ra} \dd a} \\
b(a) &= f(a)c(a)
\end{align}
Replacing
$f^\star (y)$ with the full expanded expression and plugging into the thanatological renewal equation gives:
\begin{align}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{\frac{\int_{t=0}^\infty
f(t)l(t)e^{-rt}\mu(a+y)\frac{l(t+y)}{l(t)} \dd t}{\int l(t) e^{-rt} \dd
t}}{\frac{\int_{t=0}^\infty l(t)e^{-rt} \mu(t+y)\frac{l(t+y)}{l(t)} \dd t}{\int
l(t) e^{-rt} \dd t}} \dd a \dd y
\intertext{The denominator of $c(a)$ cancels out:}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{
\int_{t=0}^\infty
f(t)l(t)e^{-rt}\mu(t+y)\frac{l(t+y)}{l(t)} \dd t}{
\int_{t=0}^\infty l(t)e^{-rt} \mu(t+y)\frac{l(t+y)}{l(t)} \dd t} \dd a \dd y
\intertext{Some $l(a)$'s also cancel out:}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{
\int_{t=0}^\infty
f(t)e^{-rt}\mu(t+y)l(t+y) \dd t}{
\int_{t=0}^\infty e^{-rt} \mu(t+y)l(t+y) \dd t} \dd a \dd y
\intertext{Now $d(a+y) = \mu (a+y)l(a+y)$, so we can cancel the inner integral
with the denominator:}
=& \int _{y=0}^\infty 
\int_{a=0}^\infty
f(a)e^{-ra}\mu(a+y)l(a+y) \dd a \dd y
\intertext{Note also that $l(a) = \int _{x=a}^\infty d(x) \dd x$ (present
livings are future deaths), which brings us back to the chronological
formulation:} =& \int_{a=0}^\infty e^{-ra}l(a)f(a)
\dd a
\end{align}
\end{appendices}
\nocite{HMD,HFD}
% --------------------------------------------------
% bibliography
\bibliographystyle{plainnat}
  \bibliography{References}

>>>>

\FloatBarrier
\section*{The thanatological renewal model}
For the renewal model that follows, we are most interested in using the kind
of fertility rates shown in Figure~\ref{fig:FySpaghetti}, and we refer to a
unisex population based on female vital rates. In the following, I
explain describe the model in four different ways: as a continuous
model, as a matrix model, with a lifecylcle graph, and intuitively by
appealing to how population pyramids and leaves (described earlier) change over
time. Assuming constant vital rates, the births for the present year are given by
\begin{align}
B(t) =& \int _0^\infty P(y,t)f^\star(y) \dd y = \int _0^\infty P(a,t)f(a) \dd a
\quad\text{,} \intertext{where $a$ indexes chronological age, $y$ indexes
thanatological age, $P(a), P(y)$ are population counts, and $f^\star(y), f(a)$ are
exact specific fertility probabilities (rates), which here are best thought of
as the fertility rates of females born to females. The thanatological integral can
be broken down back in terms of chronological age:} =& \int_{y=0}^\infty
\int_{a=0}^\infty f^\star(y)\frac{P(a,t)d(a+y)}{l(a)} \dd a \dd y \quad\text{,}
\intertext{where $d(a)$ is the continuous lifetable death distribution with
radix of 1, so $d(a)/l(0)$. We can relate the present population to past births
with $P(a,t) = B(t-a)l(a)$:}
\label{eq:ergostep1}
=& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) B(t-a)d(a+y)\dd a \dd y
\quad\text{.} \intertext{Eventually-- usually a bit quicker than is the case for
chronological age-- strong ergodicity will assert itself, and $B(t)$ will be related to
$B(t-a)$ according to a constant factor $e^{ra}$, where $r$ is the familiar
intrinsic rate of growth:}
\label{eq:ergostep2}
=& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) B(t)e^{-ra}d(a+y)\dd a \dd y
\quad\text{.}
\intertext{Divide out $B(t)$ to get back a familiar-looking renewal equation:}
\label{eq:thanoren}
1 =& \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) d(a+y)e^{-ra}\dd a \dd y
\quad\text{.}
\end{align}
Now compare this to Lotka's chronological formulation
\begin{align}
\label{eq:lotka}
1 =& \int_{a=0}^\infty f(a)l(a)e^{-ra}\dd a \quad\text{,}
\intertext{which can be expressed as well as:}
  =& \int_{a=0}^\infty \int_{y=0}^\infty f(a)d(a+y)e^{-ra}\dd y \dd a
\end{align}
and note that these are really quite similar, since $\int _{y=0}^\infty
d(a+y)\dd a = l(a)$. For intuition, notice that $l(a)$ is here split
up into pieces of $d(a)$, and imagine a 2D surface of these, where one axis is chronological age and the
other axis is thanatological age. For the chronological case \eqref{eq:lotka},
we multiply chronological age-specific fertility rates, $f(a)$ over one margin,
and for the thanatological case we multiply $f^\star(y)$ over the other margin.
There is a strong parallel here with the case of two-sex age-specific fertility rates. As in the case of
divergence between male and female single-sex models, there will always be divergence between
single-sex models under chronological versus thanatological age, unless the
starting population is stable. A proof of equivalence in the case of
$f^\star(y)$ being calculated from a stable starting population is provided
in Appendix~\ref{app:B}. Even though the modeled population stocks are in a way
commensurable, the rates used here, $f(a)$ versus $f^\star(y)$ are calculated on the basis of differently distributed denominators $E(a)$ versus $E(y)$.

@@

Once one finds $r$ from \eqref{eq:thanoren}, other familiar stable population
parameters can be calculated. For instance, we may calculate the mean
thanatological generation time, $T^\star$:
\begin{equation}
\label{eq:Ty}
 T^\star =  \frac{\int _{y=0}^\infty \int _{a=0}^\infty y e^{-ra} d(a+y)
 f^\star(y) \dd a \dd y}{\int _{y=0}^\infty \int _{a=0}^\infty e^{-ra} d(a+y)
 f^\star(y) \dd a \dd y} \quad \text{.}
\end{equation}
Literally, this is the mean of the remaining lifespans of new mothers in the
stable population. The net reproduction rate, $R_0$ is related by, e.g.,
\begin{equation}
\label{eq:R0fromTy}
R_0 = e^{r T^\star}
\end{equation}
The birth rate, $b$, is given by
\begin{align}
\label{eq:eybrate}
b =& \frac{1}{\int _{y=0}^\infty \int _{a=0}^\infty e^{-ra} d(a+y) \dd a \dd
y}\quad\text{,}
\intertext{the denominator of which is the inside of
equation~\eqref{eq:thanoren} less $f^\star(y)$. This simplifies to} =&
\frac{1}{ \int _{a=0}^\infty e^{-ra} l(a) \dd a } \quad \text{,}
\intertext{which is exactly the same as that valid for
chronological age in the Lotka-Euler setup, except that $r$ here comes from the
thanatological renewal model. If the stable age structure, $c^\star(y)$, is
known, then we may take the sum of hypothetical births over thanatological age:}
=& \int _{y=0}^\infty f^\star(y) c^\star(y) \dd y \quad{.}
\end{align}

The stable thanatological age structure, $c^\star(y)$, is the proportion of the stable population with remaining years to live $y$:
\begin{align}
\label{eq:cy}
c^\star(y) =& b \int _{a=0}^\infty e^{-ra} d(a+y) \dd a 
\intertext{or, given the stable chronological age structure and a lifetable:}
           =& \int _{a=0}^\infty c(a)d(y|a) \dd a\\ \notag \\
           =& \frac{\int _{a=0}^\infty d(a+y)e^{-ra}\dd a}{\int
           _{a=0}^\infty l(a)e^{-ra}\dd a}  \quad\text{.}
\end{align}

It has previously been shown that when $r = 0$ and $y = a$, that
$c(a)=c^\star(y)$ \citep{brouard1989mouvements, vaupel2009life}, and now we can arrive at the same
conclusion directly by equating the chronological and thanatological renewal equations. For
small non-zero magnitude values of $r$ and survival patterns typical of humans, it is also the
case that $c^\star(y|r)\approx c(a|-r)$. This could make a useful heuristic,
and the relationship is asymptotically exact as $r$ approaches zero, but the
exact relationship is to be found in equation~\eqref{} Further
stable population quantities may be estimated by similarly translating the various common definitions (e.g.,
in the glossary of \citet{coale1972growth}) to the present perspective. We will
focus on the main model rather than on these.

The thanatological renewal model is valid, but this perspective may take some
time to gestate before accepting that it is also a sound description of how
populations renew. What should one imagine under the model of thanatological
population renewal? A useful mnemonic bases itself on Figures~\ref{fig:USdecomp}
and \ref{fig:USrecomp}. In the chronological age-structured model, new
generations appear at the bottom of the pyramid, and move up one rung per year.
All age-classes are subject to attrition, which is spread out over ages and not
readily visible in the pyramid. In the thanatological \textit{leaf}, each birth cohort increments to the population over the whole range of thanatological age
according to $d(a)$, as seen in \eqref{eq:thanoren}, becoming the
shaded layers seen in Figure~\ref{fig:USrecomp}. Each horizontal step is a death
cohort, and these move one step down the pyramid each year without any decrement
(indeed incrementing due to births) until reaching the very bottom. In short,
the locations of increment and decrement, and the direction of movement (when
so visualized) are all switched. The chronological and thanatological renewal
models are almost perfectly opposite descriptions of the same process.

@@

\begin{appendices}
\section{Unique solution for thanatological $r$}
\label{app:A}
This appendix contains a brief proof that the real solution for the
intrinsic growth rate, $r$, is unique for the case of the thanatological renewal model, \eqref{eq:thanoren}, and can be proven so in
essentially the same fashion as those in existence for the Lotka-Euler model,
\eqref{eq:lotka}. This proof follows that given in
\citet{pressat1973analyse}. Define a convenience function, $I(r)$, for the
integrand of \eqref{eq:thanoren} for a given $r$ and fixed $f^\star(y)$ and $d(a)$:

\begin{equation}
I(r) = \int_{y=0}^\infty \int_{a=0}^\infty f^\star(y) d(a+y)e^{-ra}\dd a \dd y
\end{equation}
Since the death distribution function, $d(a)$, and fertility
function, $f^\star(y)$, are continuous and non-negative, $\lim_{r \to +\infty} I(r)
= 0$ and $\lim_{r \to -\infty} I(r)= \infty$. If $r_2 > r_1$, then $I(r_1) >
I(r_2)$. $I()$ is therefore a continuous and monotonically decreasing function
of $r$ with boundaries that include the value 1 of \eqref{eq:thanoren}, and
 necessarily only obtain this value once. As with the Lotka-Euler equation, there will be more complex
 conjugate solutions for $r$ in the thanatological model,
 and these are not explored in this paper.

\section{Equivalence in stability}
\label{app:B}
This appendix contains a proof that the thanatological renewal equation implies
the same intrinsic growth rate, $r$, as the Lotka-Euler model if
the starting population is stable. This is identical to claiming that the right
side of the rewal equation is equal given the same $r$. Since there is a unique
real solution, demonstrated in the previous appendix, then it is sufficient to
show that the two integral equations are equal under these conditions. The basic
relationship is:
\begin{align}
\int _{a=0}^\infty l(a)e^{-ra}f(a) \dd a &= \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra}f^\star (y) \dd a \dd y
\intertext{when}
f^\star(y) &= \frac{b^\star(y)}{c^\star(y)} \\
b^\star (y) &= \int_{a=0}^\infty b(a) \mu(a+y))\frac{l(a+y)}{l(a)} \dd a \\
c^\star (y) &= \int_{a=0}^\infty c(a) \mu(a+y))\frac{l(a+y)}{l(a)} \dd a
\intertext{Note that:}
c(a) &= \frac{l(a)r^{-ra}}{\int l(a) e^{-ra} \dd a} \\
b(a) &= f(a)c(a)
\end{align}Replacing $f^\star (y)$ with the full expanded
expression and plugging into the thanatological renewal equation gives:
\begin{align}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{\frac{\int_{t=0}^\infty
f(t)l(t)e^{-rt}\mu(a+y)\frac{l(t+y)}{l(t)} \dd t}{\int l(t) e^{-rt} \dd
t}}{\frac{\int_{t=0}^\infty l(t)e^{-rt} \mu(t+y)\frac{l(t+y)}{l(t)} \dd t}{\int
l(t) e^{-rt} \dd t}} \dd a \dd y
\intertext{The denominator of $c(a)$ cancels out:}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{
\int_{t=0}^\infty
f(t)l(t)e^{-rt}\mu(t+y)\frac{l(t+y)}{l(t)} \dd t}{
\int_{t=0}^\infty l(t)e^{-rt} \mu(t+y)\frac{l(t+y)}{l(t)} \dd t} \dd a \dd y
\intertext{Some $l(a)$'s also cancel out:}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{
\int_{t=0}^\infty
f(t)e^{-rt}\mu(t+y)l(t+y) \dd t}{
\int_{t=0}^\infty e^{-rt} \mu(t+y)l(t+y) \dd t} \dd a \dd y
\intertext{Now $d(a+y) = \mu (a+y)l(a+y)$, so we can cancel the inner integral
with the denominator:}
=& \int _{y=0}^\infty 
\int_{a=0}^\infty
f(a)e^{-ra}\mu(a+y)l(a+y) \dd a \dd y
\intertext{Note also that $l(a) = \int _{=a}^\infty d(x) \dd x$ (present
livings are future deaths), which brings us back to the chronological
formulation:} =& \int_{a=0}^\infty e^{-ra}l(a)f(a)
\dd a
\end{align}

\section{An iterative method to find the thanatological $r$}
\citet{coale1957new} proposed a fast-converging iterative approach to estimate
the intrinsic growth rate for the Lotka-Euler equation. For the thanatological
renewal model, a similar approach may be taken, with some slight
modifications to Coale's original.
The following steps can be followed to estimate $r$ from
Equation~\eqref{eq:thanoren}:

\begin{enumerate}
  \item Derive a first rough estimate of the mean remaining years of life at
  reproduction, $\widehat{T^\star}$, akin to Lotka's mean generation time, $T$.
  To start, a good-enough guess is to just assume $r=0$:
\begin{equation}
\widehat{T^\star} = \frac{\int _{y=0}^\infty \int _{a=0}^\infty y d (a+y)
f^\star(y) \dd a \dd y}{\int _{y=0}^\infty \int _{a=0}^\infty d(a+y) f^\star(y)
\dd a \dd y}
\end{equation}
  \item A first rough guess at the net reproduction rate, $R_0$ is given by
 \begin{equation}
  R_0 = \int _{y=0}^\infty \int _{a=0}^\infty d(a+y) f^\star(y) \dd a
\dd y
\end{equation}
  \item A first rough estimate of $r$, $r^0$, is given by
   \begin{equation}
   r^0 = \frac{ln(R_0)}{\widehat{T^\star}}
   \end{equation}
  \item Plug $r^0$ into Equation~\eqref{eq:thanoren} to calculate a
  residual, $\delta^0$.
  \item Use $\delta^0$ and $\widehat{T^\star}$ to calibrate the estimate of $r$
  using
  \begin{equation}
  r^{1} = r^0 + \frac{\delta^0}{\widehat{T^\star} - \frac{\delta^0}{r^0}}
  \end{equation}
  \item Repeat step (4) to to derive a new $\delta^i$, then step (5) to refine
  $r^i$, until converging on a stable $r$ after some 20 or so iterations,
  depending on the degree of precision desired ($\widehat{T^\star}$ is not updated
  in this process).
\end{enumerate}

The above procedure is more computationally efficient than minimizing the
absolute residual of Equation~\eqref{eq:thanoren} using a generic
optimizer. Alternatively, one could use the method of cumulants to estimate $r$,
bbut this would entail less precision than the above (unless one goes beyond,
say, five cumulants), and this has not yet been worked out at this 
\end{appendices}
\nocite{HMD,HFD}
-------------------------
% bibliography
\bibliographystyle{plainnat}
  \bibliography{References}
